Visualising Life Satisfaction in Countries across Europe: Individual- and Multi-Level Evidence on Cultural and Pecuniary Factors

### 2019-01-31 LaTex/Paper_fulltext.tex
Added user Manual

\section{User Manual}

>>>>

\section{User Manual}
The section is aimed at enabling the user to use the project results competently. To do so, a working model of the statistical software R is required, the following instructions are made for application within its GUI RStudio. The project is divided in 'Data preparation', "Varaible Preparation', 'Visualisation' and 'app'. The latter is the most important for the first time user as it will provide access to the reactive visualisation tool 'European Happiness Observer' that is the centrepiece of this project. First time users, may ignore the other files and can execute the whole 'app' file after installing all required packages. The visualisation tool will open itself subsequently, no further steps or data provision needed. The application has guidelines itself that should be helpful for the user in navigating through its function. Generally speaking, it offers opportunities to build mapplots for the nations and most of the regions at NUTS-1 level, other visualisations can be made using bar charts, bivariate line plots and boxplots grouped by geographical entities. Additionally, it is possible to adapt some linear regression models and display their output. Finally, raw data tables are displayed as the most unrefined view on the data. While the mapplots, bar charts and line plots can be understood quite intuitively, this paragraph is aimed at giving some guidelines for working with the 'Boxplot' and 'Regression table' elements in the application.\\

A boxplot is a visualisation that displays the distribution of data on the y-axis,  The dark bar inside the box indicates the position of the median, while the box itself indicates the value for the first and third quartile of the data. Thus the inside of the box contains 50\% of the data. The antennas around the box (also called 'whiskers') have a distance from it of 1.5 times the size of the box (distance between the third and first quartile, called interquartile range). Any data points outside the range of box and whiskers are considered outliers from the data distribution. Thus, the boxplot allows to compare the spread of data for different groups that are displayed on the x-axis. In the application it is possible, to order the groups after certain variables to see, whether the variable may have an effect on the spread of data.

\input{table1}

The regression tables provide the user with the opportunity to compile customized regression models that will give formally statistical evidence on the substance of relations in the data. The simple linear regression without multilevel elements is to be interpreted very similar to the line plot, as starting with the constant value, every increase in the independent variable changes the dependent variable by a certain amount. The number of stars next to the independent variables indicate the significance of its effect (significance as in 'being present'). The total validity of the model can be taken from the value of R$^2$ as it indicates the share of data spread in the dependent variable that is explained by the data spread in the independent variable. In table 1, this is exemplified for the model from 'Job satisfaction' on 'Life satisfaction'.\\

\input{table2}

However, by introducing multilevel elements into the regression model, the interpretation changes. Multilevel models rest on the assumption that the state of a variable as a characteristic of an individual (like life satisfaction) is expected to be influenced by structural factors. These influences can be accounted for when looking at the effect of variables that are measured on a structural level (like GDP per capita). The effect of this independent variable will then explain differences in the dependent variable that occur because there is variation between the groups, where the structural variable is measured (between-variance). It is then possible to account with variables for the individuals (like job satisfaction) exclusively for the variance inside the group (within-variance). This is exemplified in table 2. Like before, the significance of a variable is indicated by its number of stars. However, the quality of the model calculated is more difficult to assess than before. In table 2 the Akaike Information criterion (AIC) is provided for that purpose. This measure in absolute says nothing about the quality of the model but comparing several model, the one with the smallest value will be the one that explains the most.\\

Users might be interested in seeing more detailed information on variables and can do so in the codebook 'EVS\_final\_cdbk' in the project file. It is also possible to generate the codebook anew by executing the 'Visualisations file'. if users may wish to make changes to data or transformation of variables, the may do so in the files 'Data preparation' and 'Variable preparation'.
